---
title: "isye6501_hw6"
author: "Chengqi"
date: "2019/9/27"
output: word_document
---

**Question 9.1**
=================== 
Using the same crime data set uscrime.txt as in Question 8.2, apply Principal Component Analysis and then create a regression model using the first few principal components.  Specify your new model in terms of the original variables (not the principal components), and compare its quality to that of your solution to Question 8.2.  You can use the R function prcomp for PCA. (Note that to first scale the data, you can include scale. = TRUE to scale as part of the PCA function. Don’t forget that, to make a prediction for the new city, you’ll need to unscale the coefficients (i.e., do the scaling calculation in reverse)!)

**\textcolor{red}{Answer}**  

First we have to load data.
```{r}
#load data
data <- read.table('C:\\Users\\huangchengqi\\Desktop\\MS SCE\\19Fall\\ISYE6501\\hw6\\data 9.1\\uscrime.txt',header=TRUE)
str(data)
```

Now we do PCA to the original data. Since there are 15 predictors, we will have 15 principal components. Meanwhile, we need to scale the original data.
```{r}
#do PCA
PCA_Model<-prcomp(data[,1:15],scale=TRUE)
summary(PCA_Model)

#first we choose 10 pricipal components to build an original model.
data2 <- PCA_Model$x[,1:10]
pca_crime <- data.frame(cbind(data2,data[,16]))
model_1 <- lm(V11~.,pca_crime)
model_1$coefficients
summary(model_1)
```

We then use stepwise method to remove predictors from model_1 in order to find a model with lowest AICc.
```{r, include=FALSE}
stepAICc <- function (object, scope, scale = 0, direction = c("both", "backward", 
  "forward"), trace = 1, keep = NULL, steps = 1000, use.start = FALSE, 
  k = 2, ...) 
{
  mydeviance <- function(x, ...) {
    dev <- deviance(x)
    if (!is.null(dev)) 
      dev
    else MuMIn::AICc(x, k=0)
  }
  cut.string <- function(string) {
    if (length(string) > 1L) 
      string[-1L] <- paste("\n", string[-1L], sep = "")
    string
  }
  re.arrange <- function(keep) {
    namr <- names(k1 <- keep[[1L]])
    namc <- names(keep)
    nc <- length(keep)
    nr <- length(k1)
    array(unlist(keep, recursive = FALSE), c(nr, nc), list(namr, 
      namc))
  }
  step.results <- function(models, fit, object, usingCp = FALSE) {
    change <- sapply(models, "[[", "change")
    rd <- sapply(models, "[[", "deviance")
    dd <- c(NA, abs(diff(rd)))
    rdf <- sapply(models, "[[", "df.resid")
    ddf <- c(NA, abs(diff(rdf)))
    AIC <- sapply(models, "[[", "AIC")
    heading <- c("Stepwise Model Path \nAnalysis of Deviance Table", 
      "\nInitial Model:", deparse(formula(object)), "\nFinal Model:", 
      deparse(formula(fit)), "\n")
    aod <- if (usingCp) 
      data.frame(Step = change, Df = ddf, Deviance = dd, 
        `Resid. Df` = rdf, `Resid. Dev` = rd, Cp = AIC, 
        check.names = FALSE)
    else data.frame(Step = change, Df = ddf, Deviance = dd, 
      `Resid. Df` = rdf, `Resid. Dev` = rd, AIC = AIC, 
      check.names = FALSE)
    attr(aod, "heading") <- heading
    class(aod) <- c("Anova", "data.frame")
    fit$anova <- aod
    fit
  }
  Terms <- terms(object)
  object$formula <- Terms
  if (inherits(object, "lme")) 
    object$call$fixed <- Terms
  else if (inherits(object, "gls")) 
    object$call$model <- Terms
  else object$call$formula <- Terms
  if (use.start) 
    warning("'use.start' cannot be used with R's version of 'glm'")
  md <- missing(direction)
  direction <- match.arg(direction)
  backward <- direction == "both" | direction == "backward"
  forward <- direction == "both" | direction == "forward"
  if (missing(scope)) {
    fdrop <- numeric()
    fadd <- attr(Terms, "factors")
    if (md) 
      forward <- FALSE
  }
  else {
    if (is.list(scope)) {
      fdrop <- if (!is.null(fdrop <- scope$lower)) 
        attr(terms(update.formula(object, fdrop)), "factors")
      else numeric()
      fadd <- if (!is.null(fadd <- scope$upper)) 
        attr(terms(update.formula(object, fadd)), "factors")
    }
    else {
      fadd <- if (!is.null(fadd <- scope)) 
        attr(terms(update.formula(object, scope)), "factors")
      fdrop <- numeric()
    }
  }
  models <- vector("list", steps)
  if (!is.null(keep)) 
    keep.list <- vector("list", steps)
  n <- nobs(object, use.fallback = TRUE)
  fit <- object
  bAIC <- extractAIC(fit, scale, k = k, ...)
  edf <- bAIC[1L]
  bAIC <- MuMIn::AICc(fit, k=k)
  if (is.na(bAIC)) 
    stop("AIC is not defined for this model, so 'stepAIC' cannot proceed")
  if (bAIC == -Inf) 
    stop("AIC is -infinity for this model, so 'stepAIC' cannot proceed")
  nm <- 1
  Terms <- terms(fit)
  if (trace) {
    cat("Start:  AIC=", format(round(bAIC, 2)), "\n", cut.string(deparse(formula(fit))), 
      "\n\n", sep = "")
    utils::flush.console()
  }
  models[[nm]] <- list(deviance = mydeviance(fit), df.resid = n - 
    edf, change = "", AIC = bAIC)
  if (!is.null(keep)) 
    keep.list[[nm]] <- keep(fit, bAIC)
  usingCp <- FALSE
  while (steps > 0) {
    steps <- steps - 1
    AIC <- bAIC
    ffac <- attr(Terms, "factors")
    if (!is.null(sp <- attr(Terms, "specials")) && !is.null(st <- sp$strata)) 
      ffac <- ffac[-st, ]
    scope <- factor.scope(ffac, list(add = fadd, drop = fdrop))
    aod <- NULL
    change <- NULL
    if (backward && length(scope$drop)) {
      aod <- dropterm(fit, scope$drop, scale = scale, 
        trace = max(0, trace - 1), k = k, ...)
      rn <- row.names(aod)
      row.names(aod) <- c(rn[1L], paste("-", rn[-1L], 
        sep = " "))
      if (any(aod$Df == 0, na.rm = TRUE)) {
        zdf <- aod$Df == 0 & !is.na(aod$Df)
        nc <- match(c("Cp", "AIC"), names(aod))
        nc <- nc[!is.na(nc)][1L]
        ch <- abs(aod[zdf, nc] - aod[1, nc]) > 0.01
        if (any(is.finite(ch) & ch)) {
          warning("0 df terms are changing AIC")
          zdf <- zdf[!ch]
        }
        if (length(zdf) > 0L) 
          change <- rev(rownames(aod)[zdf])[1L]
      }
    }
    if (is.null(change)) {
      if (forward && length(scope$add)) {
        aodf <- addterm(fit, scope$add, scale = scale, 
          trace = max(0, trace - 1), k = k, ...)
        rn <- row.names(aodf)
        row.names(aodf) <- c(rn[1L], paste("+", rn[-1L], 
          sep = " "))
        aod <- if (is.null(aod)) 
          aodf
        else rbind(aod, aodf[-1, , drop = FALSE])
      }
      attr(aod, "heading") <- NULL
      if (is.null(aod) || ncol(aod) == 0) 
        break
      nzdf <- if (!is.null(aod$Df)) 
        aod$Df != 0 | is.na(aod$Df)
      aod <- aod[nzdf, ]
      if (is.null(aod) || ncol(aod) == 0) 
        break
      nc <- match(c("Cp", "AIC"), names(aod))
      nc <- nc[!is.na(nc)][1L]
      o <- order(aod[, nc])
      if (trace) {
        print(aod[o, ])
        utils::flush.console()
      }
      if (o[1L] == 1) 
        break
      change <- rownames(aod)[o[1L]]
    }
    usingCp <- match("Cp", names(aod), 0) > 0
    fit <- update(fit, paste("~ .", change), evaluate = FALSE)
    fit <- eval.parent(fit)
    nnew <- nobs(fit, use.fallback = TRUE)
    if (all(is.finite(c(n, nnew))) && nnew != n) 
      stop("number of rows in use has changed: remove missing values?")
    Terms <- terms(fit)
    bAIC <- extractAIC(fit, scale, k = k, ...)
    edf <- bAIC[1L]
    bAIC <- MuMIn::AICc(fit, k=k)
    if (trace) {
      cat("\nStep:  AIC=", format(round(bAIC, 2)), "\n", 
        cut.string(deparse(formula(fit))), "\n\n", sep = "")
      utils::flush.console()
    }
    if (bAIC >= AIC + 1e-07) 
      break
    nm <- nm + 1
    models[[nm]] <- list(deviance = mydeviance(fit), df.resid = n - 
      edf, change = change, AIC = bAIC)
    if (!is.null(keep)) 
      keep.list[[nm]] <- keep(fit, bAIC)
  }
  if (!is.null(keep)) 
    fit$keep <- re.arrange(keep.list[seq(nm)])
  step.results(models = models[seq(nm)], fit, object, usingCp)
}
```

Apply the Stepwise Method:
```{r}
#stepwise method
library(MuMIn)
library(MASS)
model_1 <- lm(V11~.,pca_crime)
stepAICc(model_1,direction='both', steps=1000)
```

The result shows it is better to use pc1, pc2, pc4, pc5, pc7 as predictors to build the linear regression model.

```{r}
#use the 5 mentioned predictors
data3 <- PCA_Model$x[,c(1,2,4,5,7)]
pca_crime_2 <- data.frame(cbind(data3,data[,16]))
model_2 <- lm(V6~.,pca_crime_2)
summary(model_2)
```

We need to have the coefficients of the original predictors. 
```{r}
#calculate a0 and ai
a0 <- model_2$coefficients[1]
ai <- PCA_Model$rotation[,c(1,2,4,5,7)] %*% model_2$coefficients[2:6]
a0
ai
```

However, this coefficients are calculated using the scaled data. In order to do prediction to a given city, we need to unscale the coefficients.
While scaling, we subtracted the mean and divides by the standard deviation, for each variable. We then have this equation:
ai * (x - mean)/sd+a0 = original_ai * x+original_a0.
That means:
(1) original_ai = ai/sd
(2) original_a0 = a0-ai*mean/sd, 
```{r}
#do the scale calculation in reverse
originalai <- ai/sapply(data[,1:15],sd)
originala0 <- a0 - sum(ai*sapply(data[,1:15],mean)/sapply(data[,1:15],sd))
```

Now we have a proper regression model. We can use this model to predict number of crimes in a given city. We use the data given in question 8.2, which is:
M = 14.0
So = 0
Ed = 10.0
Po1 = 12.0
Po2 = 15.5
LF = 0.640
M.F = 94.0
Pop = 150
NW = 1.1
U1 = 0.120
U2 = 3.6
Wealth = 3200
Ineq = 20.1
Prob = 0.04
Time = 39.0
```{r}
#apply pca to predict crimes using the data given in question 8.2
test_data_set <-data.frame(M = 14,So = 0, Ed = 10.0, Po1 = 12.0, Po2 = 15.5,LF = 0.64, M.F = 94.0, Pop = 150, NW = 1.1, U1 = 0.12, U2 = 3.6, Wealth = 3200, Ineq = 20.1, Prob = 0.040, Time = 39.0)
result <- as.matrix(test_data_set) %*% as.matrix(originalai)+originala0 
result
```
The result shows the number of crimes predicted is about 1357, which is reasonable.